---
title: "DS_Project"
author: "Allison Shafer, Monica Puerto, Alison Ragan"
date: "10/27/2019"
output: html_document
---

```{r setup, include=FALSE}

# call in libraries to use

suppressMessages(library(tidyverse))
suppressMessages(library(modelr))
suppressMessages(library(lubridate))
suppressMessages(library(stringr))
suppressMessages(library(purrr))
suppressMessages(library(gganimate))
suppressMessages(library(readxl))
suppressMessages(library(gifski))
suppressMessages(library(tidycensus))
suppressMessages(library(rvest))
<<<<<<< Updated upstream
=======
suppressMessages(library(tmap))
suppressMessages(library(spData))
suppressMessages(library(tigris))
suppressMessages(library(sf))

# 11/8: Added packages "tmap", "spData", "tigris"
#install.packages("tmap")
#install.packages("spData")
#install.packages("tigris")

>>>>>>> Stashed changes
```


```{r download data - eviction dataset}

# set up a data directory if it does not exist already
if(!file.exists("./data")) {dir.create("./data")}

# store urls needed - right click on excel symbol and select 
# copy link address
eviction_US_all <- c("https://eviction-lab-data-downloads.s3.amazonaws.com/US/all.csv")

# check to see if URL saved to variable
eviction_US_all

#download the csv file
download.file(eviction_US_all, destfile = "./data/eviction_US_all.csv", mode = "wb") #this takes a bit ranges from 2-5 minutes, over bad WIFI might take 10 min

# read the csv file into R and save into dataframe
eviction_US_all <- read_csv("./data/eviction_US_all.csv")

eviction_US_all

#get download speed statistics
start <- Sys.time()
end <- Sys.time()

print(end-start)
```

<<<<<<< Updated upstream
```{r rural urban}
=======
```{r import rural urban}
>>>>>>> Stashed changes

# download and load urbal/rural classifications

# set up a data directory if it does not exist already
if(!file.exists("./data")) {dir.create("./data")}

# store urls needed - right click on excel symbol and select 
# copy link address
urban_rural <- c("http://www2.census.gov/geo/docs/reference/ua/County_Rural_Lookup.xlsx")

#download the .xlsx file

download.file(urban_rural, destfile = "./data/urban_rural.xlsx", mode = "wb")

#read xlsx into R

urban_rural <- read_excel(skip = 3, "./data/urban_rural.xlsx")

names(urban_rural)

# clean data set
urban_rural <- urban_rural %>%
  rename("Percent_Rural" = "2010 Census \r\nPercent Rural",
         "GEOID" = "2015 GEOID") 

# check names were updated
names(urban_rural)

# select only needed fields
urban_rural <-urban_rural %>%
  mutate(rural_flag = ifelse(round(Percent_Rural, 3) > 50, 1, 0)) %>%
  select(GEOID, State, Percent_Rural, rural_flag) 


# view cleaned up dataset
urban_rural

```

```{r Webscrape US Labor Stats}
<<<<<<< Updated upstream
setwd("/Users/monicapuerto/Desktop/Github/eviction_lab_project")
=======
>>>>>>> Stashed changes

start <- Sys.time()
if(!file.exists("./data/labor_data")) {dir.create("./data/labor_data")}

#save url into a variable
url <- "https://www.bls.gov/lau/#tables"

#download the html content using read_html
download.file(url,destfile="./data/uslabor.html")
us_county_labor_html <- read_html("./data/uslabor.html")

#extract the xslx 
us_county_labor_html %>% 
   rvest::html_nodes("ul") %>%
        rvest::html_nodes("li") %>%
        rvest::html_nodes("a") %>%
        rvest::html_attr("href") %>%
        str_subset(".xlsx$") -> us_labor_urls

#domain
domain <- "https://www.bls.gov"

#paste domain to urls
str_c(domain,us_labor_urls) -> us_labor_urls

#only need years from 2000 to 2016
us_labor_urls[3:19] -> us_labor_2000_2016
years <- rep(2000:2016,1)

#a for loop that downloads each file
for(i in seq_along(us_labor_2000_2016)){
  download.file(us_labor_2000_2016[i],destfile = paste("./data/labor_data/",years[i],".xslx",sep=""),mode="wb")
}

#save the files pertaining to us labor
labor_files <- dir("./data/labor_data")


#create a function that downloads each url and saves it #into a dataframe
read_files <- function(x){
read_excel(path= paste("./data/labor_data/",x,sep=""),skip = 7,col_names = c("laus_code","state_fips_code","county_fips_code","county_name","year","","labor_force","employed","unemployed","unemployment_rate"),sheet = 1,na="")
}

#map the function to read each file 
map(labor_files,read_files) -> all_labor_data

#join all the US labor tables
all_labor_data %>% reduce(full_join) -> all_labor_data

#remove some rows that have NA in Year and remove empty column next to year
filter(all_labor_data,!is.na(year)) %>% select(-X__1) -> all_labor_data
 
 end<- Sys.time()

 print(end-start)
 
#creating a GEOID to join 
all_labor_data$GEOID <- str_c(all_labor_data$state_fips_code,all_labor_data$county_fips_code)
 
 
#change year to integer
all_labor_data$year <- as.integer(all_labor_data$year)
 
```

```{r make county dataset}

<<<<<<< Updated upstream
```{r data exploration, echo = false, eval = false}

#look at col names
colnames(eviction_US_all)

#look at unique values for each variable
map(eviction_US_all, unique)


```

```{r make county dataset}
#County Table joined with 2 additional variables : rural flag and unemployment rate
=======
#County Table joined with 2 additional variables: rural flag and unemployment rate
>>>>>>> Stashed changes
eviction_county <- eviction_US_all %>%
  right_join(urban_rural, key = "GEOID") %>% 
  left_join(select(all_labor_data,GEOID,year,unemployment_rate), by =c("GEOID" = "GEOID", "year"="year"))
         
eviction_county 

```

```{r make state dataset}

eviction_state <- eviction_US_all %>%
  filter(nchar(GEOID) == 2)
         
eviction_state

```

```{r make place dataset}
<<<<<<< Updated upstream

eviction_place <- eviction_US_all %>%
  filter(nchar(GEOID) == 7)
         
eviction_place

```

=======

eviction_place <- eviction_US_all %>%
  filter(nchar(GEOID) == 7)
         
eviction_place

```

```{r download county spatial data}

# set up a data directory if it does not exist already
if(!file.exists("./data")) {dir.create("./data")}

#download county shapefiles
download.file("https://www2.census.gov/geo/tiger/GENZ2018/shp/cb_2018_us_county_500k.zip", destfile = "./data/county_shp.zip")

spatial_data <- "./data/county_shp.zip"

unzip(spatial_data, overwrite = TRUE, exdir = "./data/spatial/county_shp")



county_boundaries <- st_read("./data/spatial/county_shp/cb_2018_us_county_500k.shp")
county_boundaries

```

```{r data exploration, echo = false, eval = false}

#look at col names
colnames(eviction_US_all)

#look at summary of data
summary(eviction_US_all)

summary(all_labor_data)

```


```{r EDA county - eviction rate}

# get top 100 counties with highest average evistion rate
eviction_county %>%
  group_by(GEOID) %>%
  summarise(
    avg_rate = mean(`eviction-rate`, na.rm = TRUE)
  ) %>%
  top_n(100) %>%
  arrange(desc(avg_rate))


# histogram of eviction rates in county dataset
eviction_county %>%
  group_by(`eviction-rate`) %>%
  count() %>%
  ggplot(aes(`eviction-rate`)) +
  geom_histogram()

```


>>>>>>> Stashed changes
```{r EDA state level data}

# look at dataset

names(eviction_state)

summary(eviction_state)

# look at 2016 
eviction_state %>%
  filter(year == 2016) %>%
  ggplot(aes(x = name, y = `poverty-rate`)) +
  geom_bar(stat = 'identity') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

```


```{r EDA county level data}

# look at data set
eviction_county
<<<<<<< Updated upstream

ncol(eviction_county)

summary(eviction_county)

glimpse(eviction_county)

eviction_county %>%
  group_by(GEOID) %

```
=======

ncol(eviction_county)
>>>>>>> Stashed changes

summary(eviction_county)

glimpse(eviction_county)

eviction_county %>%
  group_by(GEOID) 

```

<<<<<<< Updated upstream
```{r map attempt}
data("county_laea", package = "tidycensus")

data("state_laea", package = "tidycensus")
```
=======
```{r gganimate eviction for state}

# look at eviction rate over time per state -- create gganimate
# this needs to have better formatting for final.

eviction_state %>%
  ggplot(aes(x = `eviction-rate`, y = name)) +
  geom_point(aes(size = population, fill = `rent-burden`),
             shape = 21) +
  scale_x_log10(breaks = 2^(-1:7)*1000) +
  scale_size(range = c(1, 20), guide = FALSE) +
  labs(x = "Eviction Rate",
       y = "State") +
  transition_states(year, transition_length = 1,
                    state_length = 1)+
  ggtitle("Year showing {closest_state}",
          subtitle = "Frame {frame} of {nframes}")+
  theme_bw()
```


```{r set up spatial datasets}

# get Census TIGER shapefiles for state
state_us_geo <- tigris::states(class= "sf")


# add spatial data to the state level eviction data
eviction_by_state <- eviction_state %>%
                      left_join(us_geo, c("GEOID" = "GEOID"))

eviction_by_state

# add spatial data to the county level eviction data
eviction_by_county <- eviction_county %>%
                      left_join(county_boundaries, c("GEOID" = "GEOID"))

eviction_by_county
```
>>>>>>> Stashed changes
